---
title: "Individual1"
author: "Brad Gannon"
date: "April 16, 2019"
output: html_document
---

Load data
```{r}
library("tidyverse")
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
names(acc)
person <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors=FALSE)
```

#Part one: Accident data
```{r}
# Are there some days of the week where more accidents happen than on others?
p <- ggplot(acc, aes(x=DAY_WEEK)) + geom_histogram(color="black", fill="blue", binwidth = 0.5)
p <- p + xlab("Day of Week (1=Sunday, 7=Saturday)") + ylab("Number of Accidents") + ggtitle("Number of Accidents by Day of Week")
p
```

It appears that the rate of accidents increases on Thursday through Sunday, which would make sense.

```{r}
# What time of the day do accidents happen?
p <- ggplot(acc, aes(x=HOUR)) + geom_histogram(color="red", fill="white", binwidth=2)
p <- p + xlab("Hour") + ylab("Number of Accidents") + ggtitle("Number of Accidents by Hour")
p
```

This histogram shows that accident rates tend to increase later in the day. The bar near 100 is signifying an unknown time of the accident.

```{r}
# What is the number of accidents with at least one drunk driver?
dim(acc)
wDrunk <- filter(acc, DRUNK_DR > 0)
dim(wDrunk)
```

Based on the dimensions, it appears that 8769 out of 34247 accidents involved at least one drunk driver, which is a rate of 25.6%

#Part two: Connecting data
```{r}
# Load the person table. Identify drivers (PER_TYPE == 1) and subset on them
driver <- filter(person, PER_TYP == 1)
```
```{r}
# Join accident and driver table (work out which variable(s) to use)
# The ST_CASE variable is a unique number for each accident, will be sufficient to inner join tables
accDriver <- merge(driver, acc, by="ST_CASE")
```
```{r}
# Tally number of accidents by day of the week, hour of the day, and gender. Visualize the results!
byDay <- ggplot(accDriver, aes(x=DAY_WEEK)) + geom_histogram(color="green", fill="grey26", binwidth=1)
byDay <- byDay + xlab("Day of Week") + ylab("Number of Accidents") + ggtitle("Number of Accidents by Day of Week")
byDay <- byDay + scale_x_continuous(breaks = seq(1, 7, 1))
byDay

byHour <- ggplot(accDriver, aes(x=HOUR.x)) + geom_histogram(color="red", fill="grey26", binwidth=1)
byHour <- byHour + xlab("Hour") + ylab("Number of Accidents") + ggtitle("Number of Accidents by Hour")
byHour

byGend <- ggplot(accDriver, aes(x=SEX)) + geom_histogram(color="blue", fill="grey26", binwidth=1)
byGend <- byGend + xlab("Gender (1=Male, 2=Female)") + ylab("Number of Accidents") + ggtitle("Number of Accidents by Gender")
byGend <- byGend + scale_x_continuous(breaks = seq(0, 10, by=1))
byGend
```

#Part three: Connecting data and make maps
```{r}
library(maps)
counties <- map_data("county")

glc <- readxl::read_excel("GLC.xlsx", skip = 1) %>%
  select(c("State Name","State Code","County Code" ,"County Name"))

names(glc) <- toupper(c("region","State","County" ,"subregion"))

glc <- glc %>%
  mutate(STATE = as.integer(STATE),
         COUNTY= as.integer(COUNTY))

newdata <- acc %>%
  group_by(STATE, COUNTY) %>%
  summarise(Total= n()) %>% 
  left_join(glc, by = c("STATE", "COUNTY") ) 
names(newdata) <- tolower(names(newdata))

newdata$region <- tolower(newdata$region)
newdata$subregion <- tolower(newdata$subregion)

choro <- inner_join(counties, newdata, by = c("region","subregion"))
choro <- choro[order(choro$order), ]

ggplot(choro, aes(long, lat)) +
  geom_polygon(aes(group = group, fill = total),color = "white")

states <- map_data("state")
glc <- readxl::read_excel("GLC.xlsx", skip = 1) %>%
  select(c("State Name","State Code")) %>%
  mutate("State Code" = as.integer(`State Code`)) %>%
  unique()

names(glc) <- toupper(c("region","STATE"))

'%ni%' <- Negate('%in%')

w <- acc %>%
  filter(WEATHER %ni% c(99,98)) %>%
  #mutate(WEATHER_NAME = ifelse(WEATHER >=3,'Winter','Summer')) %>% # as.factor(WEATHER)) %>%
  group_by(STATE,WEATHER) %>%
  summarise(Total = n()) %>% 
  left_join(y = glc, by = 'STATE')
names(w) <- tolower(names(w))
w$region <- tolower(w$region)
w <- left_join(states, w, by="region")

ggplot(w, aes(long, lat)) +
  geom_polygon(aes(group = group, fill = weather),color = "white")
```