---
title: "individual2.Rmd"
author: "Seydou Yacouba"
date: "April 17, 2019"
output: html_document
---

```{r}
library(tidyverse)
theme_set(theme_light())
acc <- read.csv("https://raw.githubusercontent.com/DS202-at-ISU/labs/master/data/fars2016/accident.csv", stringsAsFactors = FALSE)
head(acc)
```
# Plot latitude and longitude on a map

```{r}
# acc %>%
#  filter(LATITUDE == 99.9999) %>% head()
```

## Map
```{r}
acc %>%
  filter(!LATITUDE > 99) %>%
  filter(!LONGITUD > 500) %>%
  ggplot(aes(y=LATITUDE, x=LONGITUD)) +
  geom_point(size=0.1)
```

# Part One: Accident Data

* Are there some days of the week more accidents happen than on others (use varible `DAY_WEEK`)

```{r}
acc %>%
  group_by(DAY_WEEK) %>%
  tally(sort=TRUE)
```

* what time of the day do accidents happen (use `variable HOUR`)?

```{r fig.width=12, fig.height=10}
acc %>%
  filter(!HOUR > 24) %>%
  group_by(HOUR) %>%
  tally() %>%
  ggplot(aes(x=HOUR, y=n)) +
  geom_col() +
  scale_x_continuous(breaks=seq(0,23))
```
__Highest fatality counts happen between noon and 10PM.__

* what is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?

```{r fig.width=12, fig.height=10}
acc %>%
  filter(DRUNK_DR > 0) %>%
  group_by(DRUNK_DR) %>%
  tally()
```

__There are `r 8474 + 243 + 3` drunk driving accidents__


# Part two: Connecting Data

## Reading person data
```{r}
persons <- read_csv("https://raw.githubusercontent.com/DS202-at-ISU/labs/master/data/fars2016/person.csv")
```

* Connect to the person table. Identify drivers (`PER_TYP` == 1, see fars manual) and subset on them.

### Subseting drivers data

```{r}
drivers <- persons %>%
  filter(PER_TYP == 1)
```

* Join accident and driver table (work out which variable(s) to use)

```{r}
print(dim(acc))
print(dim(drivers))
```
### Performing Joing
* Joining on `STATE`, `MONTH`, `HOUR`

```{r}
joined <- left_join(acc, drivers, by=c("STATE", "MONTH", "HOUR"))
```

* Tally the number of accidents by day of the week (`DAY_WEEK`), hour of the day (`HOUR`) and gender (`SEX`).

```{r}
tally_join <- joined %>%
                group_by(DAY_WEEK,HOUR,SEX) %>%
                filter(HOUR < 24) %>%
                filter(SEX < 3) %>%
                tally()
head(tally)
```

### Visualizing the results

```{r fig.width=18, fig.height=10}
days <- c("Mon", "Tue", "Wed", "Thur", "Fri", "Sat", "Sun")
sex <- c("M","F")

tally_join %>%
  ggplot(aes(x = HOUR, y = n)) +
    geom_point() + 
    facet_wrap(factor(DAY_WEEK, labels = days ) ~ factor(SEX,labels = sex), 
               scales = "free")
```
#Part three: Connecting data and make maps
```{r}
library(maps)
counties <- map_data("county")

```

```{r warning=FALSE}
glc <- readxl::read_excel("GLC.xlsx", skip = 1) %>%
  select(c("State Name","State Code","County Code" ,"County Name"))

names(glc) <- toupper(c("region","State","County" ,"subregion"))

glc <- glc %>%
  mutate(STATE = as.integer(STATE),
         COUNTY= as.integer(COUNTY))
```

```{r}
newdata <- acc %>%
  group_by(STATE, COUNTY) %>%
  summarise(Total= n()) %>% 
  left_join(glc, by = c("STATE", "COUNTY") ) 
names(newdata) <- tolower(names(newdata))
```
```{r}
newdata$region <- tolower(newdata$region)
newdata$subregion <- tolower(newdata$subregion)
```

```{r}
#counties <- counties %>% mutate(region = toupper(region))
choro <- inner_join(counties, newdata, by = c("region","subregion"))
choro <- choro[order(choro$order), ]
```

```{r}
ggplot(choro, aes(long, lat)) +
  geom_polygon(aes(group = group, fill = total),color = "white") 
```

```{r warning=FALSE}
states <- map_data("state")
glc <- readxl::read_excel("GLC.xlsx", skip = 1) %>%
  select(c("State Name","State Code")) %>%
  mutate("State Code" = as.integer(`State Code`)) %>%
  unique()

names(glc) <- toupper(c("region","STATE"))

'%ni%' <- Negate('%in%')

w <- acc %>%
  filter(WEATHER %ni% c(99,98)) %>%
  #mutate(WEATHER_NAME = ifelse(WEATHER >=3,'Winter','Summer')) %>% # as.factor(WEATHER)) %>%
  group_by(STATE,WEATHER) %>%
  summarise(Total = n()) %>% 
  left_join(y = glc, by = 'STATE')
names(w) <- tolower(names(w))
w$region <- tolower(w$region)
w <- left_join(states, w, by="region")

```
```{r}
ggplot(w, aes(long, lat)) +
  geom_polygon(aes(group = group, fill = weather),color = "white")

```





